# A primer on functional programming

What you'll have learned by the end of the chapter: writing your own functions, functional programming basics (map, reduce, anonymous functions).

## Introduction

Functional programming is a way of write programs that relies exclusively on functions. Mathematical
functions have a very property: for any given input, they ALWAYS return exactly the same output.
This is what we want to achieve with the functions that we will write. Functions that always 
return the same result are called pure, and a language that only allows writing pure functions
is called a pure functional programming language. R is not a pure functional programming language,
so we have to be careful not to write unpure function that manipulate the global state.

But what is state? Run the following code in your console:

```{r, eval = F}
ls()
```

This will list every object defined in your state. Now run the following line:

```{r, eval = F}

x <- 1

```

and then `ls()`. `x` should now be listed alongside the other objects. You just manipulated the state 
of your current R session. Now if you run something like:

```{r, eval = F}

x + 1

```

This will produce `2`. We want to avoid pipelines that depend on some definition of some variable somewhere, 
which could be subject to change, because this could mean that 2 different runs of the same pipeline could 
produce 2 different results. Notice that I used the verb `avoid` in the sentence before. This is 
sometimes not possible to avoid. Such situations have to be carefully document and controlled.

A very practical consequence of using functional programming is that loops are not used, because
loops are imperative and imperative programming is all about manipulating state. However, there are
situations where loops are more efficient than the alternative (in R at least). So we will still
learn and use them, but only when absolutely necessary. Another consequence is that our code will
live in plain text files, and not in Jupyter (or equivalent) notebooks. Not only does imperative
code have state, but notebooks themselves have state. You should avoid notebooks at all costs, even
for experimenting.

Let's first learn about actually writing functions. Read [chapter 7](https://b-rodrigues.github.io/modern_R/defining-your-own-functions.html)
of my other book. You probably don't need to spend much time reading this chapter, as you are already
familiar with most of the concepts. Just make sure you know about the syntax if you're completely new
to R, and then jump to [chapter 8](https://b-rodrigues.github.io/modern_R/functional-programming.html).
