# A package for a project

## Introduction

In this chapter we're going to develop a package that will contain a couple of functions.
Don't focus too much on what these functions do or don't do, that's not really important.
What matters is that you understand how to build your own package, and why that's useful.

Actually, we can start by why that's useful right now.

R, as you know, has many many packages. When you type something like 

```{r, eval = F}
install.packages("dplyr")
```

This install the `{dplyr}` package. The package gets downloaded from a repository called 
CRAN - The Comprehensive R Archive Network (or from one of its mirrors). Developers thus
work on their packages and once they feel the package is ready for productions they submit
it to CRAN. There are very strict rules to respect to publish on CRAN; but if the developers
respects these rules, and the package does something *non-trivial* (that's not really defined
but the idea is that your package cannot simply be a collection of your implementation
of common mathematical functions for example), it'll get published. 

CRAN is actually quite a miracle; it works really well, and it's been working well for years,
even decades, as CRAN was founded in [1997](https://stat.ethz.ch/pipermail/r-announce/1997/000001.html).
Installing packages on R is rarely frustrating, and when it is, it is rarely, if not never,
CRAN's fault (there are some packages that require your operating system to have certain
libraries or programs installed beforehand, and these can be frustrating to install, like `java`
or certain libraries used for geospatial statistics).

But while from the point of view from the user, CRAN is great, there are sometimes some
frictions between package developers and CRAN maintainers. I'll spare you the drama, but just
know that contributing to CRAN can be sometimes frustrating.

This does not concern us however, because we are going to learn how to develop a package but
we are not going to publish it to CRAN. Instead, we will be using github.com as a replacement
for CRAN. This has the advantage that we do not have to be so strict and disciplined when 
writing our package, and other users can install the package almost just as easily from
github.com than from CRAN, with the following command:

```{r, eval = f}
remotes::install_github("github_username/some_package")
```

## Getting started

## Adding functions

### Functions dependencies

## Documentation

### Documenting functions

### Documenting the package

## Testing your package
